Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2012-05-02 paper/paper.tex
Add Ewa's changes

Large-scale computational applications expressed as scientific workflows are
often grouped into ensembles of inter-related workflows. In this paper, we
address a new and important problem concerning the efficient  management of
such ensembles under budget and deadline constraints on  Infrastructure-
as-a-Service (IaaS) clouds. We discuss, develop, and assess algorithms based
on static and dynamic strategies for both task scheduling and resource
provisioning. We evaluate the algorithms via simulation using a set of
scientific workflow ensembles with a broad range of budget and deadline
parameters, taking into account uncertainties in task runtime estimations,
provisioning delays and failures. We find that the key factor determining the
performance of an algorithm is its ability to decide which workflows in an
ensemble to admit or reject for execution. Our results show that an admission
procedure based on workflow structure and estimates of task runtimes can
significantly improve the quality of solutions.

@@

Scientific workflows, usually represented as directed acyclic graphs (DAGs), are
an important class of applications that lead to challenging problems
in resource management on grid and utility computing systems. Workflows
for large computational problems are often composed of several
inter-related workflows grouped into {\em ensembles}. All the workflows in an
ensemble typically have a similar structure, but they differ in their input
data, number of tasks, and individual task sizes.

The CyberShake~\cite{Callaghan2011} is a good example of an application that
requires scientific workflow ensembles. Generating a hazard map for a region
such as Southern California requires an ensemble of many related workflows to
compute individual hazard curves. E.g. the hazard map generated by the
CyberShake group in 2009 required an ensemble of 239 workflows. In astronomy,
for example, a user might wish to use multiple Montage~\cite{Deelman2008}
workflows with different parameters to generate a set of image mosaics that
cover a given area of the sky in different wavelengths, or use an ensemble to generate
the parts of a very large mosaic. The Pegasus team is now working on a Galactic
Plane project that is focused on generating a mosaic of the entire sky. This
ensemble consists of 17 workflows, each of which contains 900 sub-workflows and
computes one ``tile'' of the full mosaic. Another ensemble example is the
Periodograms application~\cite{Vockler2011} that searches for extrasolar planets
by detecting the periodic dips in light intensity. Due to the large scale of the
input data, this application is often split up into multiple batches processed
by different workflows. Additional workflows are created to run the analysis
using different parameters. A recent analysis of data from the Kepler satellite
required three ensembles of 15 workflows.

@@

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
computational resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the changing
demands of the application, often called auto-scaling. However, giving users 
more control also requires the development of new methods for task 
scheduling and resource provisioning. The resource management decisions required 
in cloud scenarios not only have to take into account traditional performance-related 
metrics such as workflow makespan or resource utilization, but must also now consider
budget constraints, since the resources from public (commercial) clouds
usually have monetary costs associated with them~\cite{Durkee2010}.

In this paper, we aim to gain insight into resource management challenges when
executing scientific workflow ensembles on clouds. In particular, we address a
new and important problem of maximizing the number of completed workflows from
an ensemble under both budget and deadline constraints. The motivation for this
work is to answer the fundamental question of concern to a researcher: How much
computation can be completed given the limited budget and timeframe of a
research project?

The main contributions of this paper are:
\begin{itemize}
  \item we define the problem of scheduling prioritized workflows grouped
  into ensembles, under both budget and deadline constraints,
  \item we analyze and develop dynamic (on-line) strategies for task scheduling
  and resource provisioning as well as static algorithms that
  rely on the information about the workflow structure (critical paths and
  workflow levels) and estimates of task runtimes. 
  \item we analyze and develop a hybrid workflow-aware dynamic scheduling
  algorithm, together with admission procedures to predict which workflows from
  the ensemble can be completed,
  \item we evaluate these algorithms using our simulator based on
  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the application, 
  taking into account uncertainties in task runtime
  estimations, provisioning delays and failures, 
  \item we discuss the performance of the algorithms on a set of
  synthetic workflow ensembles based on important, real scientific
  applications, on a broad range of different application scenarios and
  varying constraint values.
\end{itemize}

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper. In
Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
developed. Section~\ref{sec:performance} presents test scenarios and performance
metrics, while Section~\ref{sec:results} discusses the results of simulation
studies of these algorithms. Finally, general conclusions, lessons learned, and
future work are outlined in Section~\ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling~\cite{Autoscaling} and
RightScale~\cite{RightScale}) allow adjusting the size
of resource pool based on metrics related to infrastructure and application.
A typical infrastruc\-ture-specific metric is system load, whereas
application-speci\-fic metrics include response time and length of a task or
of a request queue. It is possible to set thresholds and limits to tune the behavior
of these autoscaling systems, but no support for complex applications is provided.

Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011}) also allow to scale the cloud resource pool or to extend the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. Our approach enables us to
take advantage of workflow-aware scheduling heuristics that cannot be applied to
independent tasks.

@@

\subsection{Resource Model}

We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
where virtual machine (VM) instances are requested on-demand and are billed by
the hour, with partial hours being rounded up. Although there may be
heterogeneous VM types with different amounts of CPU, memory, disk space, and
I/O, for this paper we focus on a single VM type because we assume that for most
applications there will typically be only one or two VM types with the best
price/performance ratio for the application~\cite{Juve2009}. We also assume that
a submitted task has exclusive access to a VM instance and that there is no
preemption. We assume that there is a delay between the time that a new VM
instance is requested and when it becomes available to execute tasks. In
practice this delay is typically on the order of tens of seconds to a few
minutes, and is highly dependent upon the cloud platform and the VM image
size~\cite{Nurmi2008b}.

\subsection{Application Model}

The target applications are ensembles of scientific workflows that can be
modeled as Directed Acyclic Graphs (DAGs), where the nodes in the graph
represent computational tasks, and the edges represent data- or control-flow
dependencies between the tasks. We assume that there are runtime estimates for
each task in the workflow based on either a performance model of the
application, or historical data that can be mined. Additionally, we assume that
the runtime estimates for individual workflow tasks are not perfect, but may
vary based on a uniform distribution of $\pm p\%$.

This study uses synthetic workflows that were generated using historical data
from real applications~\cite{Bharathi2008}. The applications come from a wide
variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA
identification protocol using high-throughput technology), astronomy (Montage),
earthquake science (CyberShake), and physics (LIGO). The runtimes used to
generate the synthetic workflows are based on distributions gathered from
running real workflows. The synthetic workflows were generated using code
developed by Bharathi, et al~\cite{WorkflowGenerator}.

Although scientific workflows are often data-intensive, the algorithms described
in the next section do not currently consider the size of input and output data
when scheduling tasks. Instead it is assumed that all workflow data is stored in
a shared cloud storage system, such as Amazon S3, and that intermediate data
transfer times are included in task runtime estimates. It is also assumed that
data transfer times between the shared storage and the VMs are equal for
different VMs so that task placement decisions do not have an impact on the
runtime of the tasks.

A workflow ensemble consists of several related workflows that a user wishes to
run. Each workflow in an ensemble is given a numeric priority that indicates how
important the workflow is to the user. As such, the priorities indicate the
utility function of the user. These priorities are absolute in the sense that
completing a workflow with a given priority is more valuable (gives the user
more utility) than completing all other workflows in the ensemble with lower
priorities combined. The goal of the workflow ensemble scheduling and cloud
provisioning problem is to complete as many high-priority workflows as possible
given a fixed budget and deadline. Only workflows for which all tasks are
finished by the deadline are considered to be complete---partial results are not
usable in this model.

@@

This section describes three algorithms were developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.

@@

DPDS' provisioning procedure is based on resource utilization. DPDS starts with
a fixed number of resources calculated based on the available time and budget,
and adjusts the number of resources according to how well they are utilized by
the application.
Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to
provision so that the entire budget is consumed before the deadline:
\begin{equation}
\label{eq:static-plan}
N_{VM} = \lceil b / (d * p) \rceil
\end{equation}
%
The DPDS algorithm provisions $N_{VM}$ VMs at the start of the ensemble execution.

The DPDS algorithm periodically computes resource
utilization using the percentage of idle VMs over time, and adjusts the number
of VMs if the utilization is above or below given thresholds. Because it is
assumed that VMs are billed by the hour, DPDS only considers VMs that are
approaching their hourly billing cycle when deciding which VMs to terminate.
This dynamic provisioning algorithm is shown in Algorithm~\ref{alg:prov}.

@@

The DPDS algorithm guarantees that tasks from lower priority workflows are
always deferred when higher-priority tasks are available, but lower-priority
tasks can still occupy idle VMs when higher-priority tasks are not available.
However, because there is no preemption, long-running low-priority tasks may
delay the execution of higher-priority tasks. In addition, tasks from low
priority workflows may be executed even though there is no chance that those
workflows will be completed within the current budget and deadline.
Fig.~\ref{fig:algorithm-example}a shows an example schedule generated using
the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
backfill idle VMs when tasks from higher priority workflows are not available.

@@

The DPDS algorithm does not use any information about the structure of the
workflows in the ensemble when scheduling tasks: it looks only at the priorities
of the ready tasks when deciding which task to schedule next. It does not
consider whether a lower priority task belongs to a workflow that will never be
able to complete given the current budget and deadline. As a result, DPDS may
start lower priority tasks just to keep VMs busy that will end up delaying
higher priority tasks later on, making it less likely that higher priority
workflows will be able to finish.

In order to address this issue, the Workflow-Aware DPDS (WA-DPDS) algorithm
extends DPDS by introducing a workflow admission procedure. The admission
procedure is invoked whenever WA-DPDS sees the first task of a new workflow at
the head of the priority queue (i.e. when no other tasks from the workflow have
been scheduled yet). The admission procedure---shown in
Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining
to admit the new workflow; if there is not, then the workflow is rejected and
its tasks are removed from the queue. This algorithm compares the current cost
(consumed budget) and remaining budget, taking into account the cost of
currently running VMs, and the cost of workflows that have already been
admitted. In addition, it adds a small safety margin of \$0.10 to avoid going
over the budget.

@@

This admission procedure relies only on the total estimated resource consumption
and compares it to the remaining budget. We found that this estimation is useful
not only to prevent low-priority workflows from delaying high-priority ones, but
also to reject large and costly workflows that would overrun the budget and
admit smaller workflows that can efficiently utilize idle resources.
% in
%ensembles containing workflows of non-uniform sizes. It would also be possible
%to extend this admission procedure to check other constraints, such as whether
%the estimated critical path of the new workflow exceeds the time remaining
%until the deadline.

@@

The previous dynamic (online) algorithms make provisioning and scheduling
decisions at runtime. In comparison, the SPSS algorithm creates a provisioning
and scheduling plan before running any workflow tasks. This enables SPSS to
start only those workflows that it knows can be completed given the deadline and
budget constraints, and eliminates any waste that may be allowed by the dynamic
algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
changes in the environment (see Sections~\ref{sec:delays} and
\ref{sec:variances}).

The approach used by SPSS is to plan each workflow in the ensemble in priority
order, rejecting any workflows that would exceed the dedline or budget. Once the
plan is complete, the VMs are provisioned and tasks are executed according to
the schedules given by the plan.

@@

To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
task in the workflow, and then schedules each task so as to minimize the cost of
the task while still meeting its assigned sub-deadline. If each task can be
completed by its deadline in the least expensive way, then the cost of the
entire workflow can be minimized without exceeding the deadline. SPSS assigns
sub-deadlines to each task based on the slack time of the workflow, which is
defined as the amount of extra time that a workflow can extend its critical path
and still be completed by the ensemble deadline. For a workflow $w$, the slack
time of $w$ is:
$$ ST(w) = d - CP(w) $$
where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
deadline and must be rejected. 
%For large ensembles we expect the critical path
%of any individual workflow to be much less than the deadline.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.45\textwidth]{figures/spds-gantt}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.45\textwidth]{figures/spss-gantt}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Tasks are represented as boxes and are colored by workflow.}
  \label{fig:algorithm-example}
\end{figure}

@@

In order to evaluate the algorithms on a standard set of workflows, we created
randomized ensembles using workflows available from the workflow generator
gallery published online by Bharathi, et al~\cite{WorkflowGenerator}. The
gallery contains synthetic workflows modeled using structures and parameters
that were taken from real applications. Ensembles were created using synthetic
workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
CyberShake. For each of these applications, workflows with 50, 100, 200, 300,
400, 500, 600, 700, 800, 900 and 1000 tasks were generated. For each workflow
size, 20 different workflow instances were generated using parameters and task
runtime distributions from traces of real workflows. The total collection of
synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
workflow instances, for a total of 1100 synthetic workflows.

@@

The number of workflows in an ensemble depends on the particular application,
but we assume that ensemble sizes are on the order of between 10 and 100
workflows. This range is motivated by several reasons: First, such sizes are
typical of the real applications we have examined (see Section~\ref{sec:intro}). 
%For example, the number of
%geographical sites of interest to the users of the CyberShake application in
%the past has been on the order of 100. 
Second, smaller ensembles consisting of
just a few workflows can be aggregated into a single workflow, so there is no need to
treat them as an ensemble. Similarly, when the number of workflows grows, and
each workflow has a large number of tasks, either the deadline and budget
constraints are low enough to prevent many of the workflows from running, or the
problem of efficiently allocating them to the resources becomes similar to a
bag-of-tasks problem, which is easier to solve efficiently.

@@

In order to quantitively compare our algorithms it is necessary to define a
metric that can be used to score the performance of the different algorithms on
a given problem (ensemble, budget, and deadline). The simplest approach is to
count the number of workflows in the ensemble that each algorithm is able to
complete within the budget before the deadline, but this metric does not account
for the priority-based utility function specified by the user. Using the
counting approach, a less efficient algorithm may be able to complete a large
number of low-priority workflows by executing the smallest workflows first.

In order to account for the user's priority, we used an exponential scoring
where the score for an ensemble $e$ is:
%
$$
Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
$$
%
where $Completed(e)$ is the set of workflows in ensemble $e$ that was
completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$
such that the highest-priority workflow has $Priority(w)=0$, the next highest
workflow has $Priority(w)=1$, and so on. This exponential scoring function
gives the highest priority workflow a score that is higher than all the lower-
priority workflows combined:
%
$$
2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
$$
%
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

@@

Finally, it is interesting to see that, for Montage and CyberShake, the
relative performance of SPSS is significantly less than it is for other
applications. We attribute this to the structure of Montage and CyberShake.
The workflows for both applications are very wide relative to their height,
and both have very short-running tasks. As a result of these two
characteristics, the Montage and CyberShake applications have very short
critical paths and look more like bag-of-tasks applications, which are easier
to execute than more structured applications. DPDS and WA-DPDS are able to
pack more of the tasks into the available budget and deadline because there
are a) more choices for where to place the tasks, and b) the different choices
have a smaller impact on the algorithms' ability to execute the workflow
within the constraints. In addition, the short critical paths put SPSS at a
disadvantage. Because of the way SPSS assigns deadlines to individual tasks,
it is prevented from starting workflows late, which prevents it from packing
tasks into the idle VM slots at the end of the schedule.

\subsection{Task Granularity}
\TODO{Should we remove this section?}

In the last experiment we noted that, for Montage and CyberShake, the short
runtimes of their tasks made the dynamic algorithms perform better relative to
SPSS. In order to test this theory we adjusted the granularity of the tasks in
several Montage and CyberShake ensembles to see how this would affect the
relative performance of the algorithms. The granularity adjustment was
achieved by multiplying the runtime of each task by a fixed scaling factor.

Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
as the scaling factor is increased from 1 to 16. The figure only shows results
for the CyberShake application with pareto unsorted ensembles, but similar
results were found for other ensemble types for both Montage and CyberShake.
Each data point in the figure represents 500 simulations (5 ensembles with 10
budgets and 10 deadlines). The best scores percent was calculated the same way
as it was for Fig.~\ref{fig:distributions}.

>>>>

Large-scale applications expressed as scientific workflows are
often grouped into ensembles of inter-related workflows. In this paper, we
address a new and important problem concerning the efficient  management of
such ensembles under budget and deadline constraints on  Infrastructure-
as-a-Service (IaaS) clouds. We discuss, develop, and assess algorithms based
on static and dynamic strategies for both task scheduling and resource
provisioning. We perform the evaluation via simulation using a set of
scientific workflow ensembles with a broad range of budget and deadline
parameters, taking into account uncertainties in task runtime estimations,
provisioning delays, and failures. We find that the key factor determining the
performance of an algorithm is its ability to decide which workflows in an
ensemble to admit or reject for execution. Our results show that an admission
procedure based on workflow structure and estimates of task runtimes can
significantly improve the quality of solutions.

@@

Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs), are
an important class of applications that lead to challenging problems
in resource management on grid and utility computing systems. Workflows
for large computational problems are often composed of several
inter-related workflows grouped into {\em ensembles}.  Workflows in an
ensemble typically have a similar structure, but they differ in their input
data, number of tasks, and individual task sizes.

CyberShake~\cite{Callaghan2011} is a good example of an application that
requires scientific workflow ensembles. Generating a hazard map for a region
such as Southern California requires an ensemble of many related workflows to
compute individual hazard curves. E.g. the hazard map generated in 2009 required an ensemble of 239 workflows. In astronomy,
a user might wish to use multiple Montage~\cite{Deelman2008}
workflows with different parameters to generate a set of image mosaics that
cover a given area of the sky in different wavelengths, or use an ensemble to generate
the parts of a very large mosaic. The
ensemble for generating a mosaic of the entire sky consists of 17 workflows, each of which contains 900 sub-workflows and
computes one ``tile'' of the full mosaic. Another ensemble example is the
Periodograms application~\cite{Vockler2011} that searches for extrasolar planets
by detecting the periodic dips in light intensity. Due to the large scale of the
input data, this application is often split up into multiple batches processed
by different workflows. Additional workflows are created to run the analysis
using different parameters. A recent analysis of Kepler satellite data
required three ensembles of 15 workflows.

@@

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the changing
demands of the application, often called auto-scaling. However, giving users 
more control also requires the development of new methods for task 
scheduling and resource provisioning. Resource management decisions required 
in cloud scenarios not only have to take into account performance-related 
metrics such as workflow makespan or resource utilization, but must also consider
budget constraints, since the resources from commercia clouds
usually have monetary costs associated with them~\cite{Durkee2010}.

In this paper, we aim to gain insight into resource management challenges when
executing scientific workflow ensembles on clouds. We address a
new and important problem of maximizing the number of completed workflows from
an ensemble under both budget and deadline constraints. The motivation for this
work is to answer the fundamental question of concern to a researcher: How much
computation can be completed given the limited budget and timeframe of a
research project?

The main contributions of this paper are:
\begin{itemize}
  \item we define the problem of scheduling prioritized workflows grouped
  into ensembles, under both budget and deadline constraints,
  \item we analyze and develop dynamic (on-line) strategies for task scheduling
  and resource provisioning as well as static algorithms that
  rely on the information about the workflow structure (critical paths and
  workflow levels) and estimates of task runtimes. 
  \item we analyze and develop a hybrid workflow-aware dynamic scheduling
  algorithm, together with admission procedures to predict which workflows from
  the ensemble can be completed,
  \item we evaluate these algorithms using our simulator based on
  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the application, 
  taking into account uncertainties in task runtime
  estimations, provisioning delays, and failures, 
  \item we discuss the performance of the algorithms on a set of
  synthetic workflow ensembles based on important, real scientific
  applications, on a broad range of different application scenarios, and
  varying constraint values.
\end{itemize}

% Ewa: the following paragraph can be deleted if necessary
%The paper is organized as follows: after an analysis of related work in
%Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
%infrastructure and application model we are targeting in this paper. In
%Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
%developed. Section~\ref{sec:performance} presents test scenarios and performance
%metrics, while Section~\ref{sec:results} discusses the results of simulation
%studies of these algorithms. Finally, general conclusions, lessons learned, and
%future work are outlined in Section~\ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling~\cite{Autoscaling} and
RightScale~\cite{RightScale}) allow adjusting the size
of resource pool based on metrics related to infrastructure and application.
A typical infrastruc\-ture-specific metric is system load, whereas
application-speci\-fic metrics include response time and length of a task or
of a request queue. It is possible to set thresholds and limits to tune the behavior
of these autoscaling systems, but no support for complex applications is provided.
Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011}) also allow to scale the cloud resource pool or to extend the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. Our approach enables us to
take advantage of workflow-aware scheduling heuristics that cannot be applied to
independent tasks.

@@

\underline{Resource Model}  We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
where virtual machine (VM) instances are requested on-demand and are billed by
the hour, with partial hours being rounded up. Although there may be
heterogeneous VM types with different amounts of CPU, memory, disk space, and
I/O, for this paper we focus on a single VM type because we assume that for most
applications there will typically be only one or two VM types with the best
price/performance ratio for the application~\cite{Juve2009}. We also assume that
a submitted task has exclusive access to a VM instance and that there is no
preemption. We assume that there is a delay between the time that a new VM
instance is requested and when it becomes available to execute tasks. In
practice this delay is typically on the order of tens of seconds to a few
minutes, and is highly dependent upon the cloud platform and the VM image
size~\cite{Nurmi2008b}.

\underline{Application Model} The target applications are ensembles of scientific workflows that can be
modeled as DAGs, where the nodes in the graph
represent computational tasks, and the edges represent data- or control-flow
dependencies between the tasks. We assume that runtime estimates for the workflow tasks
 are know and  that they are not perfect, but may
vary based on a uniform distribution of $\pm p\%$.

This study uses synthetic workflows that were generated using historical data
from real applications~\cite{Bharathi2008}. The applications come from a wide
variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA
identification protocol using high-throughput technology), astronomy (Montage),
earthquake science (CyberShake), and physics (LIGO).  The synthetic workflows were generated using code
developed in~\cite{WorkflowGenerator},  with tasks runtimes based on distributions gathered from
running real workflows.

Although workflows are often data-intensive, the algorithms described
here do not currently consider the size of input and output data
when scheduling tasks. Instead, it is assumed that all workflow data is stored in
a shared cloud storage system, such as Amazon S3, and that intermediate data
transfer times are included in task runtime estimates. It is also assumed that
data transfer times between the shared storage and the VMs are equal for
different VMs so that task placement decisions do not impact the
runtime of the tasks.

We assume that each workflow in an ensemble is given a numeric priority that indicates how
important the workflow is to the user. As such, the priorities indicate the
utility function of the user. These priorities are absolute in the sense that
completing a workflow with a given priority is more valuable than completing all other workflows in the ensemble with lower
priorities combined. The goal of the workflow ensemble scheduling and cloud
provisioning problem is to complete as many high-priority workflows as possible
given a fixed budget and deadline. Only workflows for which all tasks are
finished by the deadline are considered to be complete---partial results are not
usable in this model.

@@

This section describes three algorithms developed to schedule and provision resources for ensembles of workflows on the cloud under budget and deadline constraints.

@@

DPDS' provisioning procedure is based on resource utilization. DPDS starts with
a fixed number of resources calculated based on the available time and budget,
and adjusts the number of resources according to how well they are utilized by
the application.
Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
VM in dollars $p$, it is easy to calculate the number of VMs, $N_{VM}$, to
provision so that the entire budget is consumed before the deadline:
\begin{equation}
\label{eq:static-plan}
N_{VM} = \lceil b / (d * p) \rceil
\end{equation}
%
DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then
it periodically computes resource
utilization using the percentage of idle VMs over time, and adjusts the number
of VMs if the utilization is above or below given thresholds. Because it is
assumed that VMs are billed by the hour, DPDS only considers VMs that are
approaching their hourly billing cycle when deciding which VMs to terminate.
This dynamic provisioning algorithm is shown in Algorithm~\ref{alg:prov}.

@@

DPDS guarantees that tasks from lower priority workflows are
always deferred when higher-priority tasks are available, but lower-priority
tasks can still occupy idle VMs when higher-priority tasks are not available.
Since there is no preemption, long-running low-priority tasks may
delay the execution of higher-priority tasks. In addition, tasks from low
priority workflows may be executed even though there is no chance that those
workflows will be completed within the current budget and deadline.
Fig.~\ref{fig:algorithm-example}a shows an example schedule generated using
the DPDS algorithm. The figure illustrates how tasks from lower priority workflows
backfill idle VMs when tasks from higher priority workflows are not available.

@@

DPDS does not use any information about the structure of the
workflows in the ensemble when scheduling tasks.  It does not
consider whether a lower priority task belongs to a workflow that will never be
able to complete given the current budget and deadline. As a result, DPDS may
start lower priority tasks just to keep VMs busy that will end up delaying
higher priority tasks later on, making it less likely that higher priority
workflows will be able to finish.

The Workflow-Aware DPDS (WA-DPDS) algorithm
extends DPDS by introducing a workflow admission procedure, which is invoked whenever WA-DPDS sees the first task of a new workflow at
the head of the priority queue (i.e. when no other tasks from the workflow have
been scheduled yet). The admission procedure---shown in
Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough budget remaining
to admit the new workflow; if there is not, then the workflow is rejected and
its tasks are removed from the queue. WA-DPDS compares the current cost
(consumed budget) and remaining budget, taking into account the cost of
currently running VMs, and the cost of workflows that have already been
admitted. In addition, it adds a small safety margin of \$0.10 to avoid going
over budget. We found the admission procedure useful
not only to prevent low-priority workflows from delaying high-priority ones, but
also to reject large and costly workflows that would overrun the budget and
admit smaller workflows that can efficiently utilize idle resources.
% in
%ensembles containing workflows of non-uniform sizes. It would also be possible
%to extend this admission procedure to check other constraints, such as whether
%the estimated critical path of the new workflow exceeds the time remaining
%until the deadline.

@@

@@

The previous dynamic (online) algorithms make provisioning and scheduling
decisions at runtime. By contrast, the SPSS algorithm creates a provisioning
and scheduling plan before running any workflow tasks. This enables SPSS to
start only those workflows that it knows can be completed given the deadline and
budget constraints, and eliminates any waste that may be allowed by the dynamic
algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
changes in the environment (see Sections~\ref{sec:delays} and
\ref{sec:variances}).

%The approach used by SPSS is to plan each workflow in the ensemble in priority
%order, rejecting any workflows that would exceed the dedline or budget. Once the
%plan is complete, the VMs are provisioned and tasks are executed according to
%the schedules given by the plan.

@@

To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
task in the workflow, and then schedules each task so as to minimize the cost of
the task while still meeting its assigned sub-deadline. If each task can be
completed by its deadline in the least expensive way, then the cost of the
entire workflow can be minimized without exceeding the deadline. SPSS assigns
sub-deadlines to each task based on the slack time of the workflow, which is
defined as the amount of extra time that a workflow can extend its critical path
and still be completed by the ensemble deadline. For a workflow $w$, the slack
time of $w$ is: $ ST(w) = d - CP(w) $
where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
deadline and must be rejected. 
%For large ensembles we expect the critical path
%of any individual workflow to be much less than the deadline.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.40\textwidth]{figures/spds-gantt}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.40\textwidth]{figures/spss-gantt}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Tasks are represented as boxes and are colored by workflow.}
  \label{fig:algorithm-example}
\end{figure}

@@

In order to evaluate the algorithms on a standard set of workflows, we created
randomized ensembles using workflows available from the workflow generator
gallery~\cite{WorkflowGenerator}. 
%% Ewa: maybe put url instead of citation
The
gallery contains synthetic workflows modeled using structures and parameters
that were taken from real applications. Ensembles were created using synthetic
workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
CyberShake. For each application, workflows with 50, 100, 200, 300,
400, 500, 600, 700, 800, 900 and 1000 tasks were created. For each workflow
size, 20 different workflow instances were generated using parameters and task
runtime distributions from real workflow traces. The total collection of
synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
workflow instances, for a total of 1100 synthetic workflows.

@@

The number of workflows in an ensemble depends on the particular application,
but we assume that ensemble sizes are on the order of between 10 and 100
workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
%For example, the number of
%geographical sites of interest to the users of the CyberShake application in
%the past has been on the order of 100. 
%Second, smaller ensembles consisting of
%just a few workflows can be aggregated into a single workflow, so there is no need to
%treat them as an ensemble. Similarly, when the number of workflows grows, and
%each workflow has a large number of tasks, either the deadline and budget
%constraints are low enough to prevent many of the workflows from running, or the
%problem of efficiently allocating them to the resources becomes similar to a
%bag-of-tasks problem, which is easier to solve efficiently.

@@

In order to quantitively compare our algorithms it is necessary to define a
metric that can be used to score the performance of the different algorithms on
a given problem (ensemble, budget, and deadline). The simplest approach is to
count the number of workflows in the ensemble that each algorithm is able to
complete within the budget before the deadline, but this metric does not account
for the priority-based utility function specified by the user. Using the
counting approach, a less efficient algorithm may be able to complete a large
number of low-priority workflows by executing the smallest workflows first.
In order to account for the user's priority, we used an exponential scoring defined as:
%
$$
Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}}
$$
%
where $Completed(e)$ is the set of workflows in ensemble $e$ that was
completed by the algorithm, and $Priority(w)$ is the priority of workflow $w$
such that the highest-priority workflow has $Priority(w)=0$, the next highest
workflow has $Priority(w)=1$, and so on. This exponential scoring function
gives the highest priority workflow a score that is higher than all the lower-
priority workflows combined:
%
$$
2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i}
$$
%
%This scoring is consistent with our definition of the problem, which is to
%complete as many workflows as possible, according to their priorities, given a
%set budget and deadline.

@@

Finally, it is interesting to see that, for Montage and CyberShake, the
relative performance of SPSS is significantly less than it is for other
applications. We attribute this to the structure of Montage and CyberShake.
The workflows for both applications are very wide relative to their height,
and both have very short-running tasks, resulting  in relatively short
critical paths that makes them look more like bag-of-tasks applications, which are easier
to execute than more structured applications. DPDS and WA-DPDS are able to
pack more of the tasks into the available budget and deadline because there
are a) more choices for where to place the tasks, and b) the different choices
have a smaller impact on the algorithms' ability to execute the workflow
within the constraints. In addition, the short critical paths put SPSS at a
disadvantage. Because of the way SPSS assigns deadlines to individual tasks,
it is prevented from starting workflows late, which prevents it from packing
tasks into the idle VM slots at the end of the schedule.

%\subsection{Task Granularity}
%\TODO{Should we remove this section?}
%
%In the last experiment we noted that, for Montage and CyberShake, the short
%runtimes of their tasks made the dynamic algorithms perform better relative to
%SPSS. In order to test this theory we adjusted the granularity of the tasks in
%several Montage and CyberShake ensembles to see how this would affect the
%relative performance of the algorithms. The granularity adjustment was
%achieved by multiplying the runtime of each task by a fixed scaling factor.

%Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
%as the scaling factor is increased from 1 to 16. The figure only shows results
%for the CyberShake application with pareto unsorted ensembles, but similar
%results were found for other ensemble types for both Montage and CyberShake.
%Each data point in the figure represents 500 simulations (5 ensembles with 10
%budgets and 10 deadlines). The best scores percent was calculated the same way
%as it was for Fig.~\ref{fig:distributions}.

%The figure shows that, as the scaling factor increases beyond 2, the relative
%performance of SPSS surpasses that of the dynamic algorithms. Similar results
%were found for other ensemble types for both applications. This result
%suggests that, in general, for fine-grained workflows the dynamic algorithms
%will produce better scores, and for coarse-grained workflows SPSS will produce
%better scores.
